The Categorical Structure of Profunctor Optics

### 2017-07-17 paper.tex
Some more notes

\documentclass[11pt,a4paper]{amsart}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\let\amssquare\square
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{tensor}
\usepackage[mathscr]{eucal}
\usepackage{marvosym}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

@@

\newcommand{\id}{\mathrm{id}}

@@

\newcommand{\Pastro}{\mathrm{Pastro}}

@@

\newcommand{\Act}{\mathbf{Act}}
\newcommand{\Lens}{\mathbf{Lens}}
\newcommand{\Hask}{\mathbf{Hask}}
\newcommand{\Endo}{\mathbf{Endo}}
\newcommand{\Strong}{\mathbf{Strong}}
\newcommand{\Tamb}{\mathbf{Tamb}}
\newcommand{\Point}{\mathbf{Point}}
\newcommand{\CoPoint}{\mathbf{CoPoint}}
\newcommand{\Traversable}{\mathbf{Traversable}}

\newcommand{\op}{\mathrm{op}}
\newcommand{\ob}{\mathrm{ob}\,}

\newcommand{\fget}{\mathrm{get}}
\newcommand{\fput}{\mathrm{put}}
\newcommand{\fmodify}{\mathrm{modify}}

@@

@@

\section{Introduction}
\section{Profunctor Optics}
\section{Optic Families}
\section{The Challenge of Effectful Optics}
\section{Indexed Optics}

\nocite{*}
\bibliographystyle{amsalpha}
\bibliography{optics}
\end{document}

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\let\amssquare\square
\usepackage{mathtools}
%\usepackage{stmaryrd}
%\usepackage{tensor}
\usepackage[mathscr]{eucal}
%\usepackage{marvosym}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

@@

@@

\newcommand{\Pastro}{\Phi}
%\newcommand{\Pastro}{\mathrm{Pastro}}
\newcommand{\Double}{\mathcal{D}}

@@

\newcommand{\Act}{\mathbf{Act}}
\newcommand{\PreOptic}{\mathbf{PreOptic}}
\newcommand{\Optic}{\mathbf{Optic}}
\newcommand{\Lens}{\mathbf{Lens}}
\newcommand{\Hask}{\mathbf{Hask}}
\newcommand{\Endo}{\mathbf{Endo}}
\newcommand{\Strong}{\mathbf{Strong}}
\newcommand{\Tamb}{\mathbf{Tamb}}
\newcommand{\Point}{\mathbf{Point}}
\newcommand{\CoPoint}{\mathbf{CoPoint}}
\newcommand{\Traversable}{\mathbf{Traversable}}

\newcommand{\id}{\mathrm{id}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\const}{\mathrm{const}}
\DeclareMathOperator{\ob}{ob}
\DeclareMathOperator{\copr}{copr}

\newcommand{\fget}{\mathrm{\underline{get}}}
\newcommand{\fput}{\mathrm{\underline{put}}}
\newcommand{\fmodify}{\underline{\mathrm{modify}}}

@@

\makeatletter
\providecommand{\leftsquigarrow}{%
  \mathrel{\mathpalette\reflect@squig\relax}%
}
\newcommand{\reflect@squig}[2]{%
  \reflectbox{$\m@th#1\rightsquigarrow$}%
}
\makeatother

@@

\section{Introduction}

\todo{Might have to adjust the entire thing because the action of $\M$ on $\C$ is not strictly monoidal. On the other hand, through the entire Pastro/Street paper pretends associativity/etc is strict, so maybe who cares? What we really want is just a (nonstrict) monoidal functor $\M \to [\C, \C]$ from some $\M$.}

\todo{There might be a lot more to say in terms of Mike's stuff on framed bicategories. The ``external product'' operation on profunctors seems key. This could be the mechanism behind the indexed stuff?}

\section{Optics}

\begin{proposition}
Let $\C$ be a category and $\M$ a subcategory of $[\C, \C]$ closed under composition. There is a category $\PreOptic_\M$ that has the same objects as $\C \times \C$, with homsets given by
\begin{align*}
\PreOptic_\M((S, S'), (A, A')) = \int^{M \in \M} \C(S, M A) \times \C(M A', S')
\end{align*}
\end{proposition}
If $f : S \to M A$ and $g : M A' \to S'$, we write $\langle l, r \rangle$ for their image in $\PreOptic_\M((S, S'), (A, A'))$. We will refer to $M$ as the \emph{complement} for the pair $(l, r)$.
\begin{proof}
\todo{Following the doubles paper, this should be the Kleisli category for the monad on $\Prof$ given by:
\begin{align*}
\Phi(S, S', A, A') = \int^{M \in \M} \C(S, M A) \times \C(M A', S')
\end{align*}
}

Explicitly, the composition map
\begin{align*}
\int^{M \in \M} \left(\C(S, M A) \times \C(M A', S')\right) \times \int^{M \in \M} \left( \C(R, M S) \times \C(M S', R')\right) \to \int^{M \in \M} \C(R, M A) \times \C(M A', R')
\end{align*}
of preoptics is induced by the following morphisms for every $M, N \in \M$:
\begin{align*}
&\C(S, M A) \times \C(M A', S') \times \C(R, N S) \times \C(N S', R')\\
\xrightarrow{\text{functoriality}} \quad& \C(NS, NM A) \times \C(NM A', NS') \times \C(R, N S) \times \C(N S', R')\\
\xrightarrow{\text{composition}} \quad& \C(R, N M A) \times \C(N M A', R') \\
\xrightarrow{\copr_{NM}} \quad&\int^{M \in \M} \C(R, M A) \times \C(M A', R')
\end{align*}
In other words, let $\langle l, r \rangle \in \PreOptic_\M((R, R'), (S, S'))$ and $\langle l', r' \rangle \in \PreOptic_\M((S, S'), (A, A'))$ with $M$ the complement for $(f, g)$. The composite is then: $\langle l', r' \rangle \circ \langle l, r \rangle = \langle (M l')l, r(Mr') \rangle$
\end{proof}

In \cite{Doubles}, a very similar category was considered. The \emph{double} $\Double$ of a monoidal category $(\C, \otimes, I)$ is a special case of the above, where $\M$ is chosen to be all functors of the form $X \otimes (-)$ for $X \in \C$.

\todo{Mention the game theory stuff}

For a fixed $M \in \M$, there is a map $\C(S, M A) \times \C(M A, S) \to \C(S, S)$ given by composition. This induces a map $outside : \PreOptic_\M((S, S), (A, A)) \to \C(S, S)$. By composition in the other direction we have a map $\C(S, M A) \times \C(M A, S) \to \C(M A, M A)$. Including this object $\C(M A, M A)$ into the coend $\int^{M \in \M} \C(M A, M A)$, we induce a map $inside : \PreOptic_\M((S, S), (A,A)) \to \int^{M \in \M} \C(M A, M A)$.

\begin{definition}
An \emph{optic} from $S$ to $A$ is a preoptic $p$ from $(S, S)$ to $(A, A)$ such that $outside(p) = \id_S$, and $inside(p) = \langle \id_{MA} \rangle$ for some $M \in \M$.
\end{definition}

\begin{proposition}
\label{prop-onthenose}
If $p \in \PreOptic_\M(S, A)$ is an optic, then there exists a representative $p = \langle l, r \rangle$ such that $r \circ l = \id_{MA}$ on the nose for some $M$.
\end{proposition}
\begin{proof}
\todo{This is long and painful but seems important. There may be a more general principle at work here but I couldn't see it.}

\todo{Put in an appendix?}

In fact we will prove the more general statement that for $p \in \PreOptic_\M(S, A)$, if $outside(p) = \id_S$ and $inside(p) = \langle t \rangle$ where $t : MA \to MA$ is idempotent, then we can find a representative $\langle l, r \rangle$ with $l \circ r = t$ on the nose.

There is a well known formula for the coend as the coequaliser in the diagram
\[
\begin{tikzcd}
\coprod_{M \to N} S(N, M) \ar[r,shift left=.75ex]  \ar[r,shift right=.75ex] & \coprod_{M \in \M} S(M, M) \ar[r] & \int^{M \in \M} S(M, M)
\end{tikzcd}
\]
where $S$ is some functor $\M^\op \times \M \to \E$.

In our case $\E = \Set$, and unwinding the coequaliser this states that $\int^{M \in \M} \C(M A, M A)$ is a quotient of the set $\coprod_{M \in \M} \C(M A, M A)$. Two morphisms $f : M A \to M A$ and $g : N A \to N A$ are identified when there exists a $k : N A \to M A$ and natural transformation $\phi : M \to N$ such that the diagram
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "f", swap] & N A \ar[d, "g"] \ar[dl, "k"] \\
M A \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
commutes. This gives a binary relation $f \rightsquigarrow g$ that is not likely to be symmetric or transitive. The coend is obtained by identifying all morphisms related in this way, so in all, $\langle f \rangle = \langle g \rangle$ iff there exists a zigzag $f \leftrightsquigarrow u_1 \leftrightsquigarrow \dots \leftrightsquigarrow u_n \leftrightsquigarrow g$ where the relation may apply in either direction. \todo{replace this with $f = u_1 \leftrightsquigarrow \dots u_n = g$ for clarity}

If $f \rightsquigarrow g$ then $f^2 \rightsquigarrow g^2$: this is clear by stacking the above diagram, and then taking $k' = fk = kg$. This argument extends to show that $f^n \rightsquigarrow g^n$ for any $n$.

Similarly unwinding the coequaliser for $\int^{M \in \M} \C(S, M A) \times \C(M A, S)$, two pairs $(l, r) : (S \to M A, MA \to S)$ and $(l', r') : (S \to NA, NA \to S)$ in $\coprod_{M \in \M} \C(S, M A) \times \C(M A, S)$ are identified when there exists a $\phi : M \to N$ such that
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[d, "r'"] \\
S \ar[r, equal] \ar[d, "l", swap] & S \ar[d, "l'"]  \\
MA \ar[r, "\phi_A"] & N A
\end{tikzcd}
\]
commutes. \todo{split the diagram in the middle?}

We will now prove the result by induction. Because $inside(p) = \langle t \rangle$, there is a representative $\langle l, r \rangle$ so that $\langle lr \rangle = \langle t \rangle$. There therefore must exist a finite chain of relations $lr \leftrightsquigarrow u_1 \leftrightsquigarrow \dots \leftrightsquigarrow u_n \leftrightsquigarrow t$.
Supposing that the first relation faces rightwards, we have a diagram:
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[dd, "u_1"] \ar[ddl, "k"] \\
S \ar[d,"l",swap] & \\
M A \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
Note that $r = rlr = rk\phi_A$. We therefore have the commutative diagram
\[
\begin{tikzcd}
M A \ar[r, "\phi_A"] \ar[d, "r", swap] & N A \ar[d, "rk"] \\
S \ar[r, equal] \ar[d, "l", swap] & S \ar[d, "\phi_Al"] \\
MA \ar[r, "\phi_A", swap] & N A
\end{tikzcd}
\]
where the composite of the morphisms on the right is $\phi_A l r k = \phi_A k \phi_A k = u_1^2$. Composing in the other direction we have $r k \phi_A l = rlrl = \id_S$. We have shown that whenever $rl = \id_S$ and $lr \rightsquigarrow u_1$, then there exist $l_1$ and $r_1$ so that $r_1l_1 = \id_S$, $l_1r_1 = u_1^2$ and $(l, r) \rightsquigarrow (l_1, r_1)$. A symmetric argument shows that if instead $lr \leftsquigarrow g$, there exist $l_1$ and $r_1$ so that $r_1l_1 = \id_S$, $l_1r_1 = u_1^2$ and $(l, r) \leftsquigarrow (l_1, r_1)$.

We can now inductively apply the above argument to the chain $u_1^2 \leftrightsquigarrow \dots \leftrightsquigarrow u_n^2 \leftrightsquigarrow t^2 = t$, obtained by squaring each morphism in the original chain. We eventually obtain a chain $(l, r) \leftrightsquigarrow (l_1, r_1)\leftrightsquigarrow \dots \leftrightsquigarrow (l_n, r_n) \leftrightsquigarrow (l_t, r_t)$, where $r_tl_t = \id_S$ and $l_t r_t = t$. This pair $(l_t, r_t)$ is the required representative.
\end{proof}

The above argument has a similar form to those that appear in \cite{OnTheTrace}, which considered (among other things) coends of the form $\int^{c \in \C} \C(c, Fc)$ for an endofunctor $F : \C \to \C$.

In view of this Proposition, whenever we posit a representative $\langle l, r \rangle$ of an optic, we will assume that $l$ and $r$ are mutual two-sided inverses.

\begin{proposition}
There is an subcategory $\Optic_\M$ of $\PreOptic_\M$ given by objects of the form $(X, X)$ and optics between them.
\end{proposition}
\begin{proof}
The identity preoptic $\langle \id_X, \id_X \rangle$ is clearly an optic. Suppose we have two optics $\langle l, r \rangle \in \Optic_\M(R, S)$ and $\langle l', r' \rangle \in \Optic_\M(S, A)$, where $\langle l, r \rangle$ has complement $M$. Then the composite $\langle (Ml')l, r (Mr')  \rangle$ is certainly also an optic:
\begin{align*}
(Ml')l r (Mr') &= (Ml')(Mr') = M(l'r') = \id
\intertext{and}
r (Mr')(Ml')l &= r M(r'l') l = rl = \id
\end{align*}
\end{proof}

\subsection{Examples}

\todo{For lenses, not actually the full subcategory of $[\C, \C]$, just morphisms of the form $f \times \id$. This is a disadvantage of doing it as functors and not as an action.}

\subsection{Equivalence of Optics}

One issue with the definition of optic given in the last section is that it could be difficult to determine when two optics $\langle l, r \rangle, \langle l', r' \rangle \in \Optic_\M(S, A)$ are equal. For some choices of $\M$, however, we can reduce a zigzag of relations to a single one.

\todo{This may work in general with some huge number of conditions: something like $-A$ reflects monomorphisms and preserves binary coproducts, all subobjects in $\M$ have complements.}

\begin{proposition}
Suppose all epimorphisms split in $\C$, and that $- \times A : \C \to \C$ reflects epis. Then two lenses $p = \langle l, r \rangle$ and $q = \langle l', r' \rangle \in \Optic_\times(S, A)$ with complements $M$ and $N$ respectively are equal iff there exists an morphism $\phi : M \to N$ such that $(\phi \times A)l = l'$ and $r = r' (\phi \times A)$.
\end{proposition}
\begin{proof}
The backward direction is clear: the morphism $\phi$ is a witness that $\langle l, r \rangle \rightsquigarrow \langle l', r' \rangle$, so the two optics are equal.

%note that $p$ and $q$ must have identical $\fget$ and $\fput$ functions:
%\begin{align*}
%\fget &= \pi_2 l = \pi_2 l' \\
%\fput &= r (\pi_1 l \times A) = r' (\pi_1 l' \times A)
%\end{align*}
For the forward direction, note that the map $(\pi_1 l \times A)$ is a retraction: it has the section $(r, \pi_2)$:
\begin{align*}
(\pi_1 l \times A)(r, \pi_2) = (\pi_1 l r, \pi_2) = (\pi_1, \pi_2) = \id_{S \times A}
\end{align*}
Now by assumption, $\pi_1 l : S \to M$ is an epi, so has a retraction $f : M \to S$. The claim is that $\phi = \pi_1 l' f$ is the required morphism $M \to N$.

The situation is summarised in the following diagram, which other than the dashed line, we know commutes.
\[
\begin{tikzcd}
M \times A \ar[r, "{(f, \id_A)}", dashed, bend left = 50] \ar[d, "r", swap] & S \times A \ar[d, "\fput"] \ar[r, "{(\pi_1 l', \id_A)}"] \ar[l, "{(\pi_1 l, \id_A)}", swap] & N \times A \ar[d, "r'"] \\
S \ar[r, equal] \ar[d, "l", swap] & S \ar[d, "{(\id_S, \fget)}"] \ar[r, equal] & S \ar[d, "l'"] \\
M \times A & S \times A \ar[r] \ar[l] & N \times A
\end{tikzcd}
\]
We can now calculate:
\begin{align*}
r = r(\pi_1 l f \times A) = \fput (f \times A) = r' (\pi_1 l' f \times A) = r'(\phi \times A)
\end{align*}
as required. The equation relating $l$ and $l'$ follows immediately as they are the inverses of $r$ and $r'$.
\end{proof}

The conditions of the above Proposition hold in $\Set$, so long as $A$ is not empty.

% Old attempt:
%\begin{definition}
%A functor $F : \C \to \D$ satisfies the \emph{splitting morphism condition} (SM) if, for every $f : A \to B$ in $\C$ such that $Ff$ is a split monomorphism, there exists a $g : B \to A$ such that $F(gf) = \id_{FA}$.
%
%Similarly, $F$ satisfies the \emph{splitting epimorphism condition} (SE) if for every $f : A \to B$ in $\C$ such that $Ff$ is a split epimorphism, there exists a $g : B \to A$ such that $F(fg) = \id_{FB}$.
%\end{definition}
%
%These definitions are identical to ones given in a (as far as I can tell totally unrelated) paper on triangulated categories \cite{ObjectiveTriangleFunctors}. The conditions are not too difficult to satisfy:
%
%\begin{proposition}[\todo{cite}]
%If $F$ is full then it satisfies (SM) and (SE).
%\end{proposition}
%\begin{proof}
%
%\end{proof}
%
%\begin{proposition}[\todo{cite}]
%If $F$ is faithful then it satisfies (SM) and (SE).
%\end{proposition}
%\begin{proof}
%
%\end{proof}
%
%\begin{theorem}
%Suppose we have two optics $\langle l, r \rangle$ and $\langle l', r' \rangle \in \Optic_\M(S, A)$ with complements $M$ and $N$ respectively, and suppose $\M$ has the property that the ``evaluation-at-$A$'' functor $(-A) : \M \to \C$ satisfies (SM) and (SE). Then the two optics are equal iff there exists an morphism $\phi : M \to N$ such that $\phi_A l = l'$ and $r = r' \phi_A$.
%\end{theorem}
%\begin{proof}
%The backward direction is clear: the morphism $\phi$ is a witness that $\langle l, r \rangle \rightsquigarrow \langle l', r' \rangle$, so the two optics are equal.
%
%For the forward direction, as in Proposition \ref{prop-onthenose} we induct on a zigzag \[(l, r) = (l_1, r_1) \leftrightsquigarrow \dots \leftrightsquigarrow (l_n, r_n) = (l', r'),\] where each pair $(l_i, r_i)$ has complement $M_i$. As a base case, we can take the identity $\id_M : M \to M = M_1$. Now suppose we already have a morphism $\phi : M \to M_{i-1}$ that satisfies $\phi_A l = l_{i-1}$ and $r = r_{i-1} \phi_A$. There are two cases to consider:
%
%If $(l_{i-1}, r_{i-1}) \rightsquigarrow (l_i, r_i)$, then by definition we have a $\psi : M_{i-1}\to M_i$ that satisfies the equations $\psi_A l_{i-1} = l_i$ and $r_{i-1} = r_i \psi_A$. Therefore the composite $\psi\phi$ satisfies $(\psi\phi)_A l = \psi_A l_{i-1} = l_i$ and $r = r_{i-1} \phi_A = r_i (\psi \phi)_A$, so $\psi\phi : M \to M_i$ is the required morphism.
%
%If instead $(l_{i-1}, r_{i-1}) \leftsquigarrow (l_i, r_i)$, we have a morphism in the opposite direction $\psi : M_i \to M_{i-1}$, with $\psi_A l_i = l_{i-1}$ and $r_i = r_{i-1} \psi_A$, as shown in the following commutative diagram
%\[
%\begin{tikzcd}
%M A \ar[r, "\phi_A"] \ar[d, "r", swap] & M_{i-1} A \ar[d, "r_{i-1}"] & M_i A \ar[l, "\psi_A", swap] \ar[d, "r_i"]\\
%S \ar[r, equal] \ar[d, "l", swap] & S \ar[r, equal] \ar[d, "l_{i-1}"] & S \ar[d, "l_i"] \\
%M A \ar[r, "\phi_A", swap] & M_{i-1} A & M_i A \ar[l, "\psi_A"]
%\end{tikzcd}
%\]
%To construct a morphism $M \to M_i$ we have to perform a slightly awkward switchback maneuver. First note that $\phi_A$ is a split monomorphism, as $(l r_{i-1}) \phi_A = lr = \id_{MA}$.
%
%By condition (SM), therefore, there is a morphism $\gamma : M_{i-1} \to M$ such that $\gamma_A \phi_A = \id_{MA}$. But now, the composite $\gamma_A \psi_A : M_i \to M$ is a split epimorphism:
%\begin{align*}
%(\gamma_A \psi_A)(l_i r) = \gamma_A l_{i-1} r = \gamma_A \phi_A l r = \gamma_A \phi_A = \id_{MA}
%\end{align*}
%and by condition (SE), there is a morphism $\theta : M \to M_i$ with $\gamma_A \psi_A \theta_A = \id_{MA}$. It remains to show that $\theta_A l = l_i$ and $r = r_i \theta_A$. First some intermediate steps:
%\begin{align*}
%r \gamma_A = r_{i-1} l_{i-1} r \gamma_A = r_{i-1} \phi_A l r \gamma_A = r_{i-1} \phi_A \gamma_A = 
%\end{align*}
%So finally:
%\begin{align*}
%r &= r \gamma_A \psi_A \theta_A	
%\end{align*}
%\todo{picture}
%\end{proof}

\section{Profunctor Optics}

\subsection{Profunctors}

\todo{recall definitions} We use $I$ to refer to the identity profunctor $\C(-,{=})$, and $\odot$ for horizontal profunctor composition.

\subsection{Tambara Modules}
The following section generalises definitions that first appeared in \cite{Doubles}.
\begin{definition}
Suppose $\C$ is acted on by $\M$, a subcategory of $[\C, \C]$ closed under composition, and let $P \in \Prof(\C, \C)$ be a profunctor. A \emph{Tambara module structure for $\M$} on $P$ is a natural family of maps:
\begin{align*}
\alpha_{A,B,M} : P(A,B) \to P(MA, MB)
\end{align*}
such that $\alpha_{A,B,\id_\C} = \id_{P(A,B)}$, and the following diagram commutes:
\[
\begin{tikzcd}
P(A,B) \ar[r, "\alpha_{A,B,M}"] \ar[dr, "\alpha_{A, B, MN}" below left] & P(MA, MB) \ar[d, "\alpha_{MA, MB, N}" right] \\
& P(NMA, NMB)
\end{tikzcd}
\]
\end{definition}

Note that the hom profunctor $I := \C(-, =) : \C^\op \times \C \to \Set$ has a canonical Tambara module structure for any $\M$, given by functoriality.

If $P, Q \in \Prof(\C, \C)$ are equipped with module structures $\alpha$ and $\beta$ respectively, there is a canonical module structure on $P \odot Q$. Given $M \in \M$ and $A,B \in \C$, the structure map $(\alpha \odot \beta)_{A,B,M}$ is induced by
\begin{align*}
&P(A,C) \times Q(C,B)  \\
\xrightarrow{\alpha_{A,C,M} \times \beta_{C,B,M}} \quad& P(MA, MC) \times Q(MC, MB) \\
\xrightarrow{\copr_{MC}} \quad&\int^{C \in \C} P(MA, C) \times Q(C, MB) \\
= \quad&(P \odot Q)(MA, MB)
\end{align*}

\begin{definition}
There is a category $\Tamb_\C(\M)$ of Tambara modules and natural transformations that respect the extra structure, in the sense that for any $l : P \to Q$, the diagram
\[
\begin{tikzcd}
P(A,B) \ar[r, "\alpha_{A,B,M}"] \ar[d, "l_{A,B}" left] & P(MA, MB) \ar[d, "l_{MA, MB}" right] \\
Q(A,B) \ar[r, "\beta_{A,B,M}"] & Q(MA, MB)
\end{tikzcd}
\]
commutes.
\end{definition}

This category is monoidal with respect to $\odot$ as given above, and the monoidal unit is given by $I$. There is an evident forgetful functor $U : \Tamb_\C(\M) \to \Prof(\C, \C)$ that is strong monoidal. This forgetful functor has both a left and right adjoint; particularly important for us is the left adjoint:

\begin{definition}
Let $\Pastro_\M : \Prof(\C, \C) \to \Tamb_\C(\M)$ be the functor \todo{change notation? drop $\C$ if we use the same one the whole time?}
\begin{align*}
\Pastro_\M(P) := \int^{M \in \M}  \C(-, M{=}) \odot P \odot \C(M-, {=}) 
\end{align*}
Or, in other words, 
\begin{align*}
\mathrm{Pastro}_\C(\M)(P)(A,B) := \int^{M \in \M} \int^{C,D \in \C} \C(A, MC) \times P(C,D) \times  \C(MD, B)
\end{align*}
The module structure $\alpha_{A,B,M} : \Pastro_\M(P)(A,B) \to \Pastro_\M(P)(MA, MB)
$ is given by map induced by 
\begin{align*}
&\int^{C,D \in \C} \C(A, NC) \times P(C,D) \times  \C(ND, B) \\
\xrightarrow{\text{functoriality}} \quad& \int^{C,D \in \C} \C(MA, MNC) \times P(C,D) \times  \C(MND, MB) \\
\xrightarrow{\copr_{MN}} \quad&\int^{N \in \M} \int^{C,D \in \C} \C(MA, NC) \times P(C,D) \times  \C(ND, MB) \\
= \quad&\Pastro_\M(P)(MA, MB)
\end{align*}
for all $N \in \M$.
\end{definition}

\begin{proposition}
$\Pastro_\M : \Prof(\C, \C) \to \Tamb_\C(\M)$ is left adjoint to $U : \Tamb_\C(\M) \to \Prof(\C, \C)$.
\end{proposition}
\begin{proof}
\todo{Could crunch through this manually, may be a neat way to do it using techniques from P/S}
\end{proof}

\begin{corollary}
$\Pastro_\M$ is oplax monoidal.
\end{corollary}
\begin{proof}
This follows by abstract nonsense as a left adjoint of a strong monoidal functor. \todo{reference?}
\end{proof}

\subsection{Preoptics}

\begin{definition}
For an object $A \in \C$, the \emph{exchange profunctor} $E_A$ is defined to be $\C(-, A) \times \C(A, {=})$.
\end{definition}

Given a profunctor, or indeed a Tambara module, we can evaluate it at any two objects of $\C$. This process is functorial in the choice of Tambara module, giving a functor $(U-)(A,A) : \Tamb_\C(\M) \to \Set$.

\begin{lemma}
\label{lemma-rep}
The functor $(U-)(A,A) : \Tamb_\C(\M) \to \Set$ is representable: there is a natural isomorphism
$(U-)(A,A) \cong \Tamb_\C(\M)(\Pastro_\M E_A, -)$
\end{lemma}
\begin{proof}
We have the chain of natural isomorphisms:
\begin{align*}
&(U-)(A,A) \\
\cong \;&\int_{X,Y \in \C} \Set(\C(X,A) \times \C(A,Y), (U-)(X,Y)) && \text{(by Yoneda (un)reduction twice)} \\
=\;&\int_{X,Y \in \C} \Set(E_A(X,Y), (U-)(X,Y)) && \text{(by definition)}\\
\cong \;&\Prof(E_A, U-) && \text{(natural transformations as ends)} \\
\cong \;&\Tamb_\C(\M)(\Pastro_\M E_A, -) && \text{(by adjointness)} 
\end{align*}
by the co-Yoneda lemma twice followed by the adjunction $\Pastro_\M \dashv U$.
\end{proof}

We can now show that profunctor preoptics are precisely preoptics in the ordinary sense.

\begin{proposition}
\label{prop-profunctor-optics-are-optics}
\begin{align*}
[\Tamb_\C(\M), \Set]((U-)(A,A),(U-)(S,S)) &\cong \PreOptic_\M((S, S), (A, A))
\end{align*}
\end{proposition}
\begin{proof}
We have the chain of isomorphisms:
\begin{align*}
&[\Tamb_\C(\M), \Set]((U-)(A,A),(U-)(S,S)) \\
\cong \;&[\Tamb_\C(\M), \Set](\Tamb_\C(\M)(\Pastro_\M E_A, -), (U-)(S,S))  && \text{(by Lemma \ref{lemma-rep})}\\
\cong \;&(U\Pastro_\M E_A)(S,S)  && \text{(by Yoneda)} \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times E_A(C,D) \times \C(MD,S) \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times \C(C,A) \times \C(A,D) \times \C(MD,S) \\
\cong \;&\int^{M \in \M} \C(S, MA) \times \C(MA, S)  && \text{(by co-Yoneda twice)} \\
= \;&\PreOptic_\M((S, S), (A, A))
\end{align*}
\todo{This works for $\PreOptic_\M((S, S'), (A, A'))$ just fine, fixing the definition of $E_A$.}
\end{proof}

\begin{corollary}
A profunctor preoptic is completely determined by its component at $\Pastro_\M E_A$, and furthermore, this component is determine by its value on $\langle \id_A, \id_A, \id_A, \id_A \rangle \in (U \Pastro_\M E_A)(A, A)$.
\end{corollary}
\begin{proof}
This is the content of the first two isomorphisms above.
\end{proof}

\subsection{Optics}

The next question is: can we characterise the profunctor preoptics that actually correspond to the lawful optics? The plan is to chase through the above isomorphisms to see where the lawful optics end up.

The exchange profunctor has a canonical comonoid structure. The comultiplication $\Delta : E_A \to E_A \odot E_A$ is given by the composite
\begin{align*}
&E_A(X,Y) \\
= \quad& \C(X, A) \times \C(A, Y) \\
\xrightarrow{[\pi_1, \const_{\id_A}, \const_{\id_A}, \pi_2]} \quad& \C(X, A) \times \C(A, A) \times \C(A, A) \times \C(A, Y) \\
\xrightarrow{\copr_A} \quad& \int^{C \in \C} \C(X, A) \times \C(A, C) \times \C(C, A) \times \C(A, Y) \\
= \quad& \int^{C \in \C} E_A(X,C) \times E_A(C,Y) \\
= \quad&  (E_A \odot E_A)(X,Y)
\end{align*}
and the counit $\varepsilon : E_A \to I$ simply composes the two morphisms. Because $\Pastro_\M$ is oplax monoidal, the Tambara module $\Pastro_\M E_A$ has an induced comonoid structure.

\begin{lemma}
For an object $X$ in a monoidal category $\C$, a comonoid structure $(X,\Delta,\varepsilon)$ is equivalent to a lax monoidal structure on $\C(X, -)$.
\end{lemma}
\begin{proof}
\todo{not hard}
\end{proof}

Therefore the functor $\Tamb_\C(\M)(\Pastro_\M E_A, -)$ is canonically lax monoidal, and so the isomorphic functor $(U-)(A,A)$ is also.

\begin{proposition}
Any optic $p : (U-)(A,A) \to (U-)(S,S)$ is a lax monoidal transformation with respect to Tambara module composition in the domain and the cartesian product in the codomain.
\end{proposition}
\begin{proof}
The final few isomorphisms used in Proposition \ref{prop-profunctor-optics-are-optics} are essentially window dressing. Suppose we have an element of 
\begin{align*}
(U\Pastro_\M E_A)(S,S) = \int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times E_A(C,D) \times \C(MD,S)
\end{align*}
and let $\phi : (U\Pastro_\M E_A)(S,S) \cong \PreOptic_\M((S, S), (A, A))$ be the isomorphism given in the proposition. If we have an element of $(U\Pastro_\M E_A)(S,S)$ with representative $\langle l, e_1, e_2, r \rangle$, the round-trip through the isomorphism $\phi$ yields $\langle (M e_1)l, \id_A, \id_A, r(M e_2)\rangle$. Hence we may always assume we have a representative of the form $\langle l, \id_A, \id_A, r \rangle$. This representative corresponds to an optic via the isomorphism $\phi$ iff both $rl = \id_S$ and $\langle lr\rangle = \langle \id_{NA} \rangle$ in $\int^{M \in \M} \C(M A, M A)$. Note that just as we can ``strictify'' a representative of an optic, if we have an element of $(U\Pastro_\M E_A)(S,S)$ that corresponds to an optic we may always choose a representative so that the latter identity holds on the nose.

Now, by the definition of the Yoneda isomorphism, a natural transformation \[ \Tamb_\C(\M)(\Pastro_\M E_A, -) \to (U-)(S,S) \] will correspond to an optic iff the image of the identity morphism in $(U\Pastro_\M E_A)(S, S)$ corresponds to an optic. The very first isomorphism in Proposition \ref{prop-profunctor-optics-are-optics} is given by precomposition by the isomorphism of Lemma \ref{lemma-rep}, so in all, a profunctor preoptic $p : (U-)(A, A) \to (U-)(S,S)$ is a lawful optic iff $p(\langle \id_A, \id_A, \id_A, \id_A \rangle)$ is an optic in $(U\Pastro_\M E_A)(S, S)$. Our goal is to show that this happens exactly when the natural transformation is monoidal.

Let $\langle l, \id_A, \id_A, r \rangle = p(\langle \id_A, \id_A, \id_A, \id_A \rangle)$ be a representative, where $l : S \to M A$ and $r : M A \to S$. We can calculate the component of $p$ at a Tambara module $P$ explicitly:
\begin{align*}
p_P = (UP)(l,r) \alpha_{A,A,M}
\end{align*}
where $\alpha$ is the module structure for $P$.
